```{r packages, echo=FALSE, message=FALSE}
library(data.table)
library(dplyr)
setwd("/Users/jeherrera/Box Sync/Courses/Data Science - Specialization/5 - Reproducible Research/Week 2")
```
---
title: "Peer Assigment 1 - Reproducible Research"
output: html_document
---

This the the R Markdown document for the first assignment.

I will greatly appreciate the **comments** that you may have. Thank you!

### Loading and preprocessing the data


```{r loading_and_processing_data}
data <- read.csv("./activity.csv")
data <- filter(data, !is.na(steps))
dim(data)
head(data, 20)

data$date <- as.Date(data$date, format="%Y-%m-%d")

data$year <- as.numeric(format(data$date, format="%Y"))
data$month <- as.numeric(format(data$date, format="%m"))
data$day <- as.numeric(format(data$date, format="%d"))
```

### What is mean total number of steps taken per day?

```{r first part}
par(mfrow = c(1,1))
data <- read.csv("./activity.csv")
data <- filter(data, !is.na(steps))
data_per_prueba <- select(data, steps, date)
data_per_prueba <- group_by(data, date)
var <- summarise(data_per_prueba, suma = sum(steps))
barplot(var$suma, xlab = "Dates")

#calculate the mean
mean(var$suma)
#calculate the median
median(var$suma)
```

### What is the average daily activity pattern?

```{r second part}
data <- read.csv("./activity.csv")
data <- filter(data, !is.na(steps))
data_for_interval <- select(data, steps, interval)
data_for_interval <- group_by(data_for_interval, interval)
var1 <- summarise(data_for_interval, ave = mean(steps))
plot(var1$interval, var1$ave, type="l")

#print the max in the Average
head(arrange(var1, desc(ave)), 1)[,2]
```

### Imputing missing values

```{r third part}
data_na <- read.csv("./activity.csv")
dim(data_na[is.na(data_na$steps),])[1]
#data that does not contain NAs
data_nna <- filter(data_na, !is.na(steps))
data_per_prueba1 <- select(data_nna, steps, interval)
data_per_prueba1 <- group_by(data_nna, interval)
var <- summarise(data_per_prueba1, suma = mean(steps))
colnames(data_na)
colnames(var)
var

var[var$interval==0,]   
for (i in 1:length(data_na$steps))
{
    if(is.na(data_na[i,1]))
    {
        data_na[i,1] <- as.integer(var[var$interval==data_na[i,3],2])
    }
}

data_for_last <- data_na

data_for_final <- select(data_na, steps, date)
data_for_final <- group_by(data_for_final, date)
data_for_final <- summarise(data_for_final, m = sum(steps))
barplot(data_for_final$m, xlab = "Dates")

#calculate mean
mean(data_for_final$m)

#calculate median
median(data_for_final$m)
```


### Are there differences in activity patterns between weekdays and weekends?

```{r last part}
data_for_last$date <- as.Date(data_for_last$date, format="%Y-%m-%d")
data_for_last$day_type <- (lapply(data_for_last$date, function(x){if((weekdays(x) == "Saturday") | (weekdays(x) == "Sunday")){print("weekend")} else{print("weekday")}}))
data_for_last$day_type <- unlist(data_for_last$day_type)

sum_weekday <- select(data_for_last[data_for_last$day_type=="weekday",], steps, interval)

sum_weekend <- select(data_for_last[data_for_last$day_type=="weekend",], steps, interval)

#group_by(sum_weekday, interval, ave = mean(steps))
par(mfrow=c(2,1))
sum_weekday <- group_by(sum_weekday, interval)
sum_weekday <- summarise(sum_weekday, ave = mean(steps))
plot(sum_weekday$interval,sum_weekday$ave, type="l", main="weekday", xlab = "", ylab = "")

sum_weekend <- group_by(sum_weekend, interval)
sum_weekend <- summarise(sum_weekend, ave = mean(steps))
plot(sum_weekend$interval,sum_weekend$ave, type="l", main="weekend", xlab = "interval", ylab = "steps")

```